---
title: 'DATA CLEANING: BATCH1 - SSU SECTION A'
author: "Njambi"
date: "January 26, 2024"
output: html_document
#encoding: UTF-8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## Loading necessary libraries

```{r, include=FALSE}

library(tidyverse) # includes the dplyr, ggplot2, tidyr, readr, stringr, forcats, tibble packages
library(readxl)
library(Hmisc)
library(labelled)
library(magrittr)
library(datawizard)
library(gtsummary)
library(gt)
library(kableExtra)
library(flextable)
library(splitstackshape)

#library(ggplot2)
#library(tidyr)
#library(tibble)
#library(dplyr)
#library(stringr)

```


## Set the working directory

```{r, include=FALSE}
# Check the current working directory

getwd()
```


```{r, include=FALSE}
# Set working directory (Not necessary since the current wd is the correct wd)

setwd("C:\\Users\\awanjau\\OneDrive - CGIAR\\Alliance\\Alliance_2023\\iShamba_SSU\\SSU_Weather_SurveyAnalysis")
```

## Loading data

```{r}
# Read a csv file

ssu_sec_a <- read_csv("ssu_sec_a.csv") # Use this code when the csv file is in the set working directory

#display portion of the data

#View(ssu_sec_c)
```


## Understanding data at hand (Screening)


```{r}
# Check the dimension of the data

dim(ssu_sec_a)
```


### Generate a codebook for the data

Let's now get an overview of all variables and some more information about each variable (e.g. labels, values or value range, frequencies, missing values).  
We can achieve this by using data_codebook().


```{r, include=FALSE}

# Codebook 
# Requires package datawizard

data_codebook(ssu_sec_a)

# 34 of 40 variables shown implies that there could be empty columns in the data.

```

## Check empty columns

```{r}

# Requires package datawizard

empty_columns(ssu_sec_a)

# There are 6 empty columns in the data

```


## Drop empty columns


```{r}

# ssu_sec_a <- remove_empty_columns(ssu_sec_a)
# 
# dim(ssu_sec_a)
# 
# # There are 34 columns now.
```




## Check duplicate columns

```{r}

duplicate_columns <- ssu_sec_a[, duplicated(names(ssu_sec_a))]

#ssu_seca_clean[, empty_columns(ssu_seca_clean)]

# Print the names of duplicate columns

print(names(duplicate_columns))

# There are no duplicate columns.

```



## Check empty rows

```{r}

# Requires package datawizard

empty_rows(ssu_sec_a)

# There are no empty rows in the data.


```



# Check for duplicate rows


```{r}

duplicate_rows <- duplicated(ssu_sec_a) | duplicated(ssu_sec_a, fromLast = TRUE)

# Display the duplicate rows

print(ssu_sec_a[duplicate_rows, ])

# There are no duplicate rows

```




```{r, include=FALSE}
# Obtain the column names

names(ssu_sec_a)

# Column A6 was dropped

```



### Renaming the columns

The column names are too long for analysis.  
For ease of subsequent analysis, we will conduct this part of data cleaning process as early as now, before the data cleaning section.  

To create column names and labels, we create new column names, then use the original (long) names as column labels.  

                                                                                                                                                                                                          


```{r}

ssu_sec_a <- ssu_sec_a %>%
  rename(
    index = "index",
    respondentID = "E1. Respondent ID",
    watchedForecast = "A1. Have you watched or received any weather forecast or agro-advisory in the past 12 months?",
    interestedInForecast = "A2.1 Are you interested in receiving weather forecast or agro-advisory?",
    preferredAdvisory = "A2.2. What kind of advisory would you want to receive?",
    otherAdvisory = "77=Other (specify)...6",
    preferredChannel = "A3. What is your preferred channel for receiving weather forecasts or agro-advisory information?",
    otherChannel = "77=Other (please specify)",
    purposeForInfo = "A4. What is the main purpose for which you require weather information?",
    otherPurposeForInfo = "77=Other (specify)...10",
    importantWeatherElement = "A5. What is the MOST important element of the weather forecast that you would like to receive?",
    otherImportantWeatherElement = "77=Other (specify)...12",
    notInterestedReason = "A6. What is the primary reason for not being interested in receiving weather forecasts or agro-advisory?",
    otherNotInterestedReason = "77=Other (specify)...14",
    sourcesOfForecasts = "A7. What are the sources of weather forecasts or agro-advisories you have received in the past 12 months?",
    otherSourcesOfForecasts =  "77=Other (specify)...16",
    watchedShambaShapeUp = "Has watched Shamba shape-up weather news segment?",
    watchingDuration =  "A8. How long have you been watching the Shamba Shape Up Weather News Segment?\r\r\n\r\r\nPlease remind the farmer that this is the segment when a map of Kenya is displayed with weather information during the shamba shape up program.",
    watchWithOthers = "A9.1. Who do you usually watch the Shamba Shape-Up news segment with?",
    householdMembersWatching = "A9.2. On average, how many members of your household do you watch the Shamba Shape Up weather news segment with?",
    watchFrequency = "A10. On a scale of 1 to 5, with 5 being 'always', how often do you watch the Shamba Shape-Up weather news segment?",
    watchPlatforms = "A11. On which platform(s) do you watch or receive weather forecast information or advisory?",
    otherWatchPlatforms = "77=Other (specify)...23",
    preferredWatchPlatform = "A12. From the above platforms, which is your MOST preferred way to watch or receive weather forecast information or advisory?",
    otherPreferredWatchPlatform = "77=Other (specify)...25",
    mostImportantForecastElement = "A13. What is your MOST important element of the weather forecast to you?",
    otherImportantForecastElement = "77=Other (specify)...27",
    timelyAdvisory = "A14. Was the weather forecast information/advisory timely to guide your farming activities?",
    accuracyRating = "A15. How would you rate the accuracy of the weather forecast?",
    changedPractices = "A16. Have you changed your practices/decisions following the advisory you received?",
    changedPracticeDetails = "A17. How did you change your practice/decision?",
    otherChangedPracticeDetails = "77=Other (specify)...32",
    FarmingActivityChangesNoticed = "A18.1. Did you notice any change in your farming activities after making changes in your farming practice following the forecast/advisory?",
    positiveChangesNoticed = "A18.2. What kind of positive change did you notice?",
    otherPositiveChangesNoticed = "77=Other (specify)....35",
    negativeChangesNoticed = "A19.1. Have you noticed negative effects after making changes in your farming practice following the forecast/advisory?",
    negativeChangeDetails = "A19.2. Please specify the kind of negative effects.",
    reasonsForNoChanges = "A20. What are the main reasons why you did not make any changes, despite receiving the advisory?",
    otherReasonsForNoChanges = "77=Other (specify)....39",
    matchedWithExperience = "A21. Did the weather forecast information or advisory match with your experience/knowledge?",
    suggestionsForImprovement = "A22. What are your suggestions to improve weather forecasts or agro-advisory services?"              
  )

```



Column names are now short and standard.



## Codedebook 2

```{r}

# Codebook 
# Requires package datawizard

gt(data_codebook(ssu_sec_a))


```




## Clean all categorical variables

+ Remove "[0-9=]" from all categorical variables. (Except watchingDuration).  
+ Remove [0-9=] from watchingDuration
+ Remove "." at the end from all categorical variables except "watchingDuration".  
+ Change "watchedShambaShapeUp" from upper to title case.  


```{r}

ssu_seca_clean <- ssu_sec_a %>%
  mutate(

    watchedForecast = str_remove_all(watchedForecast, "[0-9=]|\\."),
    interestedInForecast = str_remove_all(interestedInForecast, "[0-9=]|\\."),
    preferredAdvisory = str_remove_all(preferredAdvisory, "[0-9=]|\\."),
    preferredChannel = str_remove_all(preferredChannel, "[0-9=]|\\."),
    purposeForInfo = str_remove_all(purposeForInfo, "[0-9=]|\\."),
    importantWeatherElement = str_remove_all(importantWeatherElement, "[0-9=]|\\."),
    sourcesOfForecasts = str_remove_all(sourcesOfForecasts, "[0-9=]|\\."),
    watchedShambaShapeUp = str_remove_all(watchedShambaShapeUp, "[0-9=]"),
    watchedShambaShapeUp =str_to_title(watchedShambaShapeUp),
    watchingDuration = str_remove_all(watchingDuration, "\\d+="),
    watchWithOthers = str_remove_all(watchWithOthers, "[0-9=]|\\."),
    watchFrequency = str_remove_all(watchFrequency, "[0-9=]|\\."),
    watchPlatforms	 = str_remove_all(watchPlatforms, "[0-9=]|\\."),
    preferredWatchPlatform = str_remove_all(preferredWatchPlatform, "[0-9=]|\\."),
    mostImportantForecastElement = str_remove_all(mostImportantForecastElement, "[0-9=]|\\."),
    timelyAdvisory = str_remove_all(timelyAdvisory, "[0-9=]|\\."),
    accuracyRating = str_remove_all(accuracyRating, "[0-9=]|\\."),
    changedPractices = str_remove_all(changedPractices, "[0-9=]|\\."),
    changedPracticeDetails	 = str_remove_all(changedPracticeDetails, "[0-9=]|\\."),
    FarmingActivityChangesNoticed = str_remove_all(FarmingActivityChangesNoticed, "[0-9=]|\\."),
    positiveChangesNoticed = str_remove_all(positiveChangesNoticed, "[0-9=]|\\."),
    negativeChangesNoticed = str_remove_all(negativeChangesNoticed, "[0-9=]|\\."),
    reasonsForNoChanges = str_remove_all(reasonsForNoChanges, "[0-9=]|\\."),
    matchedWithExperience = str_remove_all(matchedWithExperience, "[0-9=]|\\.")

  )



```




```{r, include=FALSE}
# columns_to_modify <- c(
#   watchedForecast, interestedInForecast, preferredAdvisory, preferredChannel, purposeForInfo, importantWeatherElement,
#   sourcesOfForecasts, watchedShambaShapeUp, watchWithOthers, watchFrequency, watchPlatforms,
#   preferredWatchPlatform, mostImportantForecastElement, timelyAdvisory, accuracyRating, changedPractices, 
#   changedPracticeDetails, FarmingActivityChangesNoticed, positiveChangesNoticed, negativeChangesNoticed,
#   reasonsForNoChanges, matchedWithExperience
# )
# 
# any(!colnames(ssu_sec_a) %in% columns_to_modify)
# 
# missing_columns <- setdiff(columns_to_modify, colnames(ssu_sec_a))
# missing_columns
# 
# ssu_seca_clean <- ssu_sec_a %>%
#   mutate(across(all_of(columns_to_modify), 
#                 ~ str_remove_all(., "[0-9=]|\\.")))

  
```




## Clean "other columns"



```{r}
dim(ssu_seca_clean)

# 106  41


```



```{r}

# Check blank columns

blank_cols <- ssu_seca_clean[, empty_columns(ssu_seca_clean)]

#sum(is.na(blank_cols))  # 636 === 6 columns  (636/106)


```


The data has 6 empty columns which will be kept for merging and appending purposes



## Obtain non-empty column names starting with "other"

```{r}

ssu_seca_clean %>%     # 106  41
  select(-empty_columns(ssu_seca_clean)) %>%  # 106  35  # Filter out blank columns
  select(starts_with("other", ignore.case = TRUE)) %>%
names()
```

## Clean each of the "other" columns separately


```{r, include=FALSE}
attach(ssu_seca_clean)

```


```{r}
unique(otherChannel)

# All unique values are initials and capitalized
```
 All unique values are initials and capitalized
 

```{r}
unique(otherSourcesOfForecasts)
```

+ iShamba has different spellings.  
+ SMS has different spellings.  
+ Could "SMS from iShamba" and "SMS-iShamba" be the same as "iShamba SMS"?



```{r}
ssu_seca_clean <- ssu_seca_clean %>%
  mutate(
    otherSourcesOfForecasts = str_replace_all(otherSourcesOfForecasts, "(?i)ishamba", "iShamba"),
    otherSourcesOfForecasts = str_replace_all(otherSourcesOfForecasts, "(?i)SMS", "SMS"),
    otherSourcesOfForecasts = str_replace_all(otherSourcesOfForecasts, "SMS-iShamba", "iShamba SMS"),
    otherSourcesOfForecasts = str_replace_all(otherSourcesOfForecasts, "SMS from iShamba", "iShamba SMS"),
    otherSourcesOfForecasts = str_squish(otherSourcesOfForecasts)
  )
  

# (?!) makes matching case insensitive

```



```{r}
unique(ssu_seca_clean$otherSourcesOfForecasts)
```



```{r}
unique(otherWatchPlatforms)
```

+ SMS has different spellings. 
+ Radio has different spellings. 
+ Google starts with lower case.  
+ Newspaper starts with lower case.
+ No consistent spacing between responses.  





```{r}
ssu_seca_clean <- ssu_seca_clean %>%
  mutate(
    otherWatchPlatforms = str_replace_all(otherWatchPlatforms, ",", ", "),
    otherWatchPlatforms = str_replace_all(otherWatchPlatforms, "(?i)sms", "SMS"),
    otherWatchPlatforms = str_replace_all(otherWatchPlatforms, "(?i)radio", "Radio"),
    otherWatchPlatforms = str_replace_all(otherWatchPlatforms, "(?i)ishamba", "iShamba"),
    otherWatchPlatforms = str_replace_all(otherWatchPlatforms, "(?i)google", "Google"),
    otherWatchPlatforms = str_replace_all(otherWatchPlatforms, "(?i)newspaper", "Newspaper"),
    otherWatchPlatforms = str_squish(otherWatchPlatforms)
  )
 



# (?!) makes matching case insensitive

```


```{r}
unique(ssu_seca_clean$otherWatchPlatforms)
```


```{r}
unique(otherPreferredWatchPlatform)
```


+ SMS has different spellings.  
+ Magazine starts with lower case.  




```{r}

ssu_seca_clean <- ssu_seca_clean %>%
  mutate(
    otherPreferredWatchPlatform = str_replace_all(otherPreferredWatchPlatform, "(?i)ishamba", "iShamba"),
    otherPreferredWatchPlatform = str_replace_all(otherPreferredWatchPlatform, "(?i)SMS", "SMS"),
    otherPreferredWatchPlatform = str_replace_all(otherPreferredWatchPlatform, "(?i)magazine", "Magazine"),
    otherPreferredWatchPlatform = str_squish(otherPreferredWatchPlatform)
  )
  

```


```{r}
unique(ssu_seca_clean$otherPreferredWatchPlatform)
```



```{r}
unique(otherImportantForecastElement)

```
+ Replace All with all the choices.  



```{r}

ssu_seca_clean <- ssu_seca_clean %>%
  mutate(otherImportantForecastElement = str_replace_all(otherImportantForecastElement, "All", "Start of rain, Likelihood of rain, Amount of rainfall, Flood, Temperature, Drought, Storm, Frost"))
  

```



```{r}
unique(ssu_seca_clean$otherImportantForecastElement)

```



```{r}

unique(otherChangedPracticeDetails)
```

+ Change all sentences to sentence case.
+ Remove extra spaces.  


```{r}

ssu_seca_clean <- ssu_seca_clean %>%
  mutate(
    otherChangedPracticeDetails =str_to_sentence(otherChangedPracticeDetails),
    otherChangedPracticeDetails = str_squish(otherChangedPracticeDetails)
  )
  
```



```{r}

unique(ssu_seca_clean$otherChangedPracticeDetails)

```


```{r}
unique(otherPositiveChangesNoticed)
```


+ Remove "harvest".  
+ Change all responses to sentence case.  
+ Replace Water harvesting...No longer have to buy coz I buy.  
+ Remove extra spaces.  




```{r}

ssu_seca_clean <- ssu_seca_clean %>%
  mutate(
    
    otherPositiveChangesNoticed =str_remove_all (otherPositiveChangesNoticed, "\\bharvest\\b"),
    otherPositiveChangesNoticed = str_replace_all(otherPositiveChangesNoticed, 
                                                  "Water harvesting...No longer have to buy coz I buy",
                                                  "I no longer have to buy water because I harvest water"),
    otherPositiveChangesNoticed = str_to_sentence(otherPositiveChangesNoticed),
    otherPositiveChangesNoticed = str_squish(otherPositiveChangesNoticed)
    
  )
  

# A regular expression called \\b makes sure that only isolated occurrences of the word "harvest" are changed, 
  # not phrases that contain other words.


```



```{r}
unique(ssu_seca_clean$otherPositiveChangesNoticed)
```


## Replace "Other (Specify)" with reponses from columns starting with "other"


```{r}

ssu_seca_clean <- ssu_seca_clean %>%
  mutate(
    
    preferredChannel = str_replace_all(preferredChannel, "Other \\(please specify\\)", replacement = otherChannel),
    sourcesOfForecasts = str_replace_all(sourcesOfForecasts, "Other \\(specify\\)", 
                                         replacement = otherSourcesOfForecasts),
    watchPlatforms = str_replace_all(watchPlatforms, "Other \\(specify\\)", replacement = otherWatchPlatforms),
    preferredWatchPlatform = str_replace_all(preferredWatchPlatform, "Other \\(specify\\)",
                                             replacement = otherPreferredWatchPlatform),
    mostImportantForecastElement = str_replace_all(mostImportantForecastElement, "Other \\(specify\\)",
                                                replacement = otherImportantForecastElement),
    changedPracticeDetails = str_replace_all(changedPracticeDetails, "Other \\(specify\\)",
                                                replacement = otherChangedPracticeDetails),
    positiveChangesNoticed = str_replace_all(positiveChangesNoticed, "Other \\(specify\\)", 
                                               replacement = otherPositiveChangesNoticed)
    
  )
 
```






## Clean open-ended questions



```{r, include=FALSE}
attach(ssu_seca_clean)
```


```{r}
unique(negativeChangeDetails)
```

+ Change case to sentence case.  
+ Replace "aboutr" with about.  
+ Remove period from "Sometimes changing of the crop is a new practice and challenging to do." for consistency. 
+ Remove comma from "I lack enough knowledge on organic farming,".
+ Remove extra spaces.  


```{r}

ssu_seca_clean <- ssu_seca_clean %>%
  mutate(
    negativeChangeDetails = str_to_sentence(negativeChangeDetails),
    negativeChangeDetails = str_replace(negativeChangeDetails, "aboutr", replacement = "about"),
    negativeChangeDetails = str_remove_all(negativeChangeDetails, "\\.$|\\,$"),
    negativeChangeDetails = str_squish(negativeChangeDetails)
    
  )
 
```



```{r}
unique(ssu_seca_clean$negativeChangeDetails)
```









```{r}
unique(suggestionsForImprovement)
```

+ Change case to sentence case. (Note the multiple responses separated by comma - 32, 62, 63)  
+ Remove "None" (case insensitive).    
+ Remove "n?a", "N/A", "n/a", "N/a", "n/A", "No suggestions".  
+ Remove periods at the end of the responses for consistency.  
+ Remove extra spaces.




```{r}

#patterns_to_remove <- c("(?i)none", "(?i)No suggestions", "\\.$", "n?a", "N/A", "n/a", "N/a", "n/A")

ssu_seca_clean <- ssu_seca_clean %>%
  mutate(
    #suggestionsForImprovement = str_to_sentence(suggestionsForImprovement),
    suggestionsForImprovement = str_remove_all(suggestionsForImprovement, 
                                      "(?i)none|(?i)No suggestions|\\.$|n\\?a|N/A|n/a|N/a|n/A"),
    suggestionsForImprovement = str_squish(suggestionsForImprovement)
    
  )
 
```




```{r}
unique(ssu_seca_clean$suggestionsForImprovement)
```






















### Column labels

After modifying the data, we can now assign labels to all the column names.


```{r}
 # Define column labels

 variable_labels <- c(

   index = "Index",
    respondentID = "E1. Respondent ID",
    watchedForecast = "A1. Have you watched or received any weather forecast or agro-advisory in the past 12 months?",
    interestedInForecast = "A2.1 Are you interested in receiving weather forecast or agro-advisory?",
    preferredAdvisory = "A2.2. What kind of advisory would you want to receive?",
    otherAdvisory = "77=Other (specify)...6",
    preferredChannel = "A3. What is your preferred channel for receiving weather forecasts or agro-advisory information?",
    otherChannel = "77=Other (please specify)",
    purposeForInfo = "A4. What is the main purpose for which you require weather information?",
    otherPurposeForInfo = "77=Other (specify)...10",
    importantWeatherElement = "A5. What is the MOST important element of the weather forecast that you would like to receive?",
    otherImportantWeatherElement = "77=Other (specify)...12",
    notInterestedReason = "A6. What is the primary reason for not being interested in receiving weather forecasts or agro-advisory?",
    otherNotInterestedReason = "77=Other (specify)...14",
    sourcesOfForecasts = "A7. What are the sources of weather forecasts or agro-advisories you have received in the past 12 months?",
    otherSourcesOfForecasts =  "77=Other (specify)...16",
    watchedShambaShapeUp = "Has watched Shamba shape-up weather news segment?",
    watchingDuration =  "A8. How long have you been watching the Shamba Shape Up Weather News Segment?\r\r\n\r\r\nPlease remind the farmer that this is the segment when a map of Kenya is displayed with weather information during the shamba shape up program.",
    watchWithOthers = "A9.1. Who do you usually watch the Shamba Shape-Up news segment with?",
    householdMembersWatching = "A9.2. On average, how many members of your household do you watch the Shamba Shape Up weather news segment with?",
    watchFrequency = "A10. On a scale of 1 to 5, with 5 being 'always', how often do you watch the Shamba Shape-Up weather news segment?",
    watchPlatforms = "A11. On which platform(s) do you watch or receive weather forecast information or advisory?",
    otherWatchPlatforms = "77=Other (specify)...23",
    preferredWatchPlatform = "A12. From the above platforms, which is your MOST preferred way to watch or receive weather forecast information or advisory?",
    otherPreferredWatchPlatform = "77=Other (specify)...25",
    mostImportantForecastElement = "A13. What is your MOST important element of the weather forecast to you?",
    otherImportantForecastElement = "77=Other (specify)...27",
    timelyAdvisory = "A14. Was the weather forecast information/advisory timely to guide your farming activities?",
    accuracyRating = "A15. How would you rate the accuracy of the weather forecast?",
    changedPractices = "A16. Have you changed your practices/decisions following the advisory you received?",
    changedPracticeDetails = "A17. How did you change your practice/decision?",
    otherChangedPracticeDetails = "77=Other (specify)...32",
    FarmingActivityChangesNoticed = "A18.1. Did you notice any change in your farming activities after making changes in your farming practice following the forecast/advisory?",
    positiveChangesNoticed = "A18.2. What kind of positive change did you notice?",
    otherPositiveChangesNoticed = "77=Other (specify)....35",
    negativeChangesNoticed = "A19.1. Have you noticed negative effects after making changes in your farming practice following the forecast/advisory?",
    negativeChangeDetails = "A19.2. Please specify the kind of negative effects.",
    reasonsForNoChanges = "A20. What are the main reasons why you did not make any changes, despite receiving the advisory?",
    otherReasonsForNoChanges = "77=Other (specify)....39",
    matchedWithExperience = "A21. Did the weather forecast information or advisory match with your experience/knowledge?",
    suggestionsForImprovement = "A22. What are your suggestions to improve weather forecasts or agro-advisory services?"  
   
     )

```


```{r}
# Set variable/column labels
 
    ssu_seca_clean <- ssu_seca_clean %>%
      set_variable_labels(.labels = variable_labels) # Requires package labelled

```


Using the function View() to open the new data in a spreadsheet reveals the column names and labels.

```{r}

View(ssu_seca_clean)

```



```{r}

ssu_sec_a_arch <- ssu_seca_clean

# Drop un needed columns and save clean data

# ssu_sec_e_arch <- ssu_sece_clean %>%
#   select(c(1:2, 5:6, 9, 12:13))

# HINT: Passing View() to a pipeline produces NULL data structure.
# Avoid the function within a pipeline

```


## Generate a data dictionary

Generate a data dictionnary and search for variables with `look_for()`

Requires package labelled.



```{r}
# Data dictionary


 ssu_B1sec_a_dict <- look_for(ssu_sec_a_arch, labels = TRUE, values = TRUE) 
# 
# # Fomart table
# 
# #flextable(data_dict)
# 
# #gt(data_dict)
# 
 #kable(ssu_B1sec_a_dict)
gt(ssu_B1sec_a_dict)

```



## Export data to CSV

## long


```{r}

#write.csv(ssu_B1sec_e_dict, "ssuB1_sec_e_arch.csv", row.names = FALSE)

```


## wide


```{r}

#write.csv(ssu_B1sec_e_dict, "ssuB1_sec_e_arch.csv", row.names = FALSE)

```



