---
title: 'DATA CLEANING: BATCH1 - SSU SECTION C'
author: "Njambi"
date: "January 23, 2024"
output: html_document
#encoding: UTF-8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## Loading necessary libraries

```{r, include=FALSE}

library(tidyverse) # includes the dplyr, ggplot2, tidyr, readr, stringr, forcats, tibble packages
library(readxl)
library(Hmisc)
library(labelled)
library(magrittr)
library(datawizard)
library(gtsummary)
library(gt)
library(kableExtra)
library(flextable)
library(splitstackshape)

#library(ggplot2)
#library(tidyr)
#library(tibble)
#library(dplyr)
#library(stringr)

```


## Set the working directory

```{r}
# Check the current working directory

getwd()
```


```{r}
# Set working directory (Not necessary since the current wd is the correct wd)

setwd("C:\\Users\\awanjau\\OneDrive - CGIAR\\Alliance\\Alliance_2023\\iShamba_SSU\\SSU_Weather_SurveyAnalysis")
```

## Loading data

```{r}
# Read a csv file

ssu_sec_c <- read_csv("ssu_sec_c.csv") # Use this code when the csv file is in the set working directory

#display portion of the data

#View(ssu_sec_c)
```


## Understanding data at hand (Screening)


```{r}
# Check the dimension of the data

dim(ssu_sec_c)
```


```{r}
# Obatain the column names

names(ssu_sec_c)
```



### Renaming the columns

The column names are too long for analysis.  
For ease of subsequent analysis, we will conduct this part of data cleaning process as early as now, before the data cleaning section.  

To create column names and labels, we create new column names, then use the original (long) names as column labels.  


```{r}
# Create new column names

colnames(ssu_sec_c) <- c(
  "index", "id", "after_wf_mitigation", "after_wf_mitigation_o",  "current_mitigation_yn",  "current_mitigation",  "current_mitigation_o"
  )

# Display a part of the data

head(ssu_sec_c)

```


Column names are now short and standard.



```{r}
# Check for duplicate rows

duplicate_rows <- duplicated(ssu_sec_c) | duplicated(ssu_sec_c, fromLast = TRUE)

# Display the duplicate rows

print(ssu_sec_c[duplicate_rows, ])

# There are no duplicate rows

```


### Generate a codebook for the data

Let's now get an overview of all variables and some more information about each variable (e.g. labels, values or value range, frequencies, missing values).  
We can achieve this by using data_codebook().


```{r}

# Codebook 
# Requires package datawizard

data_codebook(ssu_sec_c)

```


### Clean Other (Specify)

```{r}

unique(ssu_sec_c$after_wf_mitigation_o)



```



```{r}

#select(ends_with("_o"))

ssu_sec_c <- ssu_sec_c %>%
  mutate(after_wf_mitigation_o = str_to_sentence(after_wf_mitigation_o))  %>% # Capitalize only the first letter of the sentence
  mutate(after_wf_mitigation_o = str_remove_all(after_wf_mitigation_o, "\\.")) %>% # Remove all periods at the end of the sentences
  mutate(after_wf_mitigation_o = str_squish(after_wf_mitigation_o)) # Remove all extra spaces

```



```{r}

unique(ssu_sec_c$after_wf_mitigation_o)

```


### Replace "Other (Specify)" in after_wf_mitigation with after_wf_mitigation_o


```{r}
gt(data.frame(table(ssu_sec_c$after_wf_mitigation)))
```



```{r}

ssu_sec_c <- ssu_sec_c %>%
  mutate(after_wf_mitigation = str_replace_all(after_wf_mitigation, "Other \\(specify\\)",
                                               replacement = ssu_sec_c$after_wf_mitigation_o))
ssu_sec_c
```



```{r}

gt(data.frame(table(ssu_sec_c$after_wf_mitigation)))

```



### Display the unique values of after_wf_mitigation variable

```{r, include=FALSE}

# This code is similar to using ssu_sece$_ with every variable name

attach(ssu_sec_c)

```



```{r}

# Unique values of county

unique(after_wf_mitigation)

```

 


# Separate the multiple responses of after_wf_mitigation into rows

```{r}

names(ssu_sec_c)
```



```{r}

#R.Version()

View(ssu_sec_c)

dim(ssu_sec_c)
```



```{r}

ssu_secc_clean1 <- ssu_sec_c %>%
  #select(-ends_with("_o")) %>%
  mutate(after_wf_mitigation = str_replace_all(after_wf_mitigation, "[0-9=]+", replacement = "")) %>%
  mutate(after_wf_mitigation = str_remove_all(after_wf_mitigation, "\\.")) %>%
  filter(!is.na(after_wf_mitigation))  #%>%   # Remove rows with NAs
  
  
  #ssu_secc_clean1 <- splitstackshape::cSplit(ssu_secc_clean1, "after_wf_mitigation", sep = ",", direction = "wide")
  
  

#View(ssu_secc_clean1)
  
  
dim(ssu_secc_clean1)  # 104 5
 
```



```{r}

# ssu_secc_clean1 <- ssu_secc_clean1 %>%
#   separate_rows(after_wf_mitigation, sep = ", ") %>%
#   mutate(value = "Yes") %>%
#   pivot_wider(names_from = after_wf_mitigation, values_from = value, values_fill = "No")
#   
```



```{r}

dim(ssu_secc_clean1)

# long = 126  6
# wide = 104  19

View(ssu_secc_clean1)


```






```{r}

#View(ssu_sec_c)

ssu_secc_clean_long <- ssu_sec_c %>%
  select(-ends_with("_o"))%>%
  mutate(after_wf_mitigation = str_replace_all(after_wf_mitigation, "[0-9=]+", replacement = "")) %>%
  mutate(after_wf_mitigation = str_remove_all(after_wf_mitigation, "\\.")) %>%
  separate_longer_delim(after_wf_mitigation, delim = ", ") %>%
  filter(!is.na(after_wf_mitigation))  %>%  # Remove rows with NAs
  mutate(levels = "Yes", prefx = "c1") #%>%
  
# Concatenate prefx and after_wf_mitigation

ssu_secc_clean_long <- ssu_secc_clean_long %>%
  mutate(after_wf_mitigation = paste(ssu_secc_clean_long$prefx, after_wf_mitigation, sep = "_")) %>%
  select(-7)  # Drop the prefx column
  
# spread (to wide format)

  ssu_secc_clean_wide <- tidyr::spread(ssu_secc_clean_long, after_wf_mitigation, levels, fill = "No") 
  
  

#pivot_wider(names_from = after_wf_mitigation, values_from = values values_fill = "No")


dim(ssu_secc_clean_long)  # wide 126   7   

dim(ssu_secc_clean_wide)  # wide 104   19  
  

```


























```{r}

View(after_wf_mitigation)

after_wf_mitigation1 <- after_wf_mitigation %>%
  str_replace_all("[0-9=]+", replacement = "") %>%
  str_remove_all("\\.") 

  data_separated <- tibble(after_wf_mitigation1) %>%
  separate_rows(sep = ", ") %>%
    filter(!is.na(after_wf_mitigation))  %>%  # Remove rows with NAs
    mutate(values = "Yes") %>%  # Add a column with value 1
    spread(after_wf_mitigation1, values, fill = "No")


  # Spread to create binary columns
result <- spread(data_separated, mitigation_column, value, fill = 0)
  

unique(after_wf_mitigation1)


```









```{r}



# Separate the multiple responses into rows

after_wf_mitigation_long <- ssu_sec_c %>%
  select(c(id,after_wf_mitigation)) %>%
  separate_rows(after_wf_mitigation, sep = ", ") %>% 
  filter(!is.na(after_wf_mitigation))  # Remove rows with NAs

# Check the dimension of the new data

dim(after_wf_mitigation_long)

#View(after_wf_mitigation_long)

# Number of rows increased from 106 to 126

```


# Create a binary indicator variable

```{r}

# Create a binary indicator variable

after_wf_mitigation_long <- after_wf_mitigation_long %>%
  mutate(after_wf_mitigation_yesno = "Yes") # %>%
  #spread(responses, selected, fill = 0)  # Spread into wide format

# Check the dimension of the new data

dim(after_wf_mitigation_long)

```


## Separate after_wf_mitigation column into values and labels columns


```{r}

# Check unique values of after_wf_mitigation

unique(after_wf_mitigation_long$after_wf_mitigation)




after_wf_mitigation_long <- after_wf_mitigation_long %>%
  separate(
    col = after_wf_mitigation, 
    into = c("after_wf_mitigation_values", "after_wf_mitigation_labels"),
    sep = "=",
    convert = TRUE,
    remove = FALSE 
           ) 


# Remove periods at the end of after_wf_mitigation_labels column

unique(after_wf_mitigation_long$after_wf_mitigation_labels) # Unique values of the column


after_wf_mitigation_long$after_wf_mitigation_labels <- gsub("\\.$", "", after_wf_mitigation_long$after_wf_mitigation_labels)




# Convert after_wf_mitigation to factor



# Define after_wf_mitigation levels

after_wf_mitigation_levels <- c("Got insurance for my crops", "Got insurance for my livestock", "Applied for credit", "Planted smaller area", "Planted larger area", "Did not plant/did not establish crop on the farm", "Increased the number of animals", "Reduced number of livestock", "Did not keep animals", "Took up an alternative non-agricultural activity", "I didn't use any risk mitigation measures", "Other (specify)")


# Mutate after_wf_mitigation

after_wf_mitigation_long %>%
  mutate(
  after_wf_mitigation_labels = factor(
    after_wf_mitigation_labels, 
    #levels = after_wf_mitigation_levels
    levels(intersect(unique(after_wf_mitigation_long$after_wf_mitigation), after_wf_mitigation_levels))
    )  
  )

head(after_wf_mitigation_long)

dim(after_wf_mitigation_long)

```



# Convert the after_wf_mitigation_long data from long to wide



```{r}

after_wf_mitigation_long <- after_wf_mitigation_long %>%
  mutate(unique_id = row_number())

names(after_wf_mitigation_long)
```




```{r}

after_wf_mitigation_wide <- after_wf_mitigation_long %>%
  distinct(unique_id, .keep_all = TRUE) %>%
  select(unique_id, id, after_wf_mitigation_labels, after_wf_mitigation_yesno) %>%
  pivot_wider(
  id_cols = c(unique_id),
  names_from = after_wf_mitigation_labels, 
  values_from = after_wf_mitigation_yesno,
  values_fill = list(after_wf_mitigation_yesno  = "No")
  )

dim(after_wf_mitigation_wide)

View(after_wf_mitigation_wide)
```























## Observations (Diagnosis)

#### The data

+ Most of the variables have missing values.  
+ Some columns have multiple responses separated with a comma.
+ Data types are not correctly formated, eg the categorical variables are foratted as characters instead of factors.
+ Data has 106 rows and 6 columns.  
+ The column names are too long for analysis.  
+ There are no duplicate rows in the data.  
  


#### after_wf_mitigation

+ Has multiple responses in one column.  
+ Has 2 missing values.  
+ Upon separating by '=' sign, the reulsting labels column has values with period at the end



































## Data cleaning activities (Treatment)


#### The data

1. .
2. .
3. .  
4. .   
5. .
 

#### after_wf_mitigation

1. .
2. .
3. .
4. ".













### Column labels

After modifying the data, we can now assign labels to all the column names.


```{r}
 # Define column labels

variable_labels <- c(
  "index" = "Index",
  "id" = "E1..Respondent.ID",
  "gender" = "E2..Gender",
  "gender_values" = "Gender codes",
  "gender_labels" = "Gender categories",
  "county" = "E3..Which.county.do.you.reside.in.",
  "age" = "E4..How.old.are.you.",
  "age_values" = "Age codes",
  "age_labels" = "Age categories",
  "loe" = "E5..What.is.the.highest.level.of.formal.education.you.have.completed.",
  "loe_values" = "loe codes",
  "loe_labels" = "loe categories",
  "hhsize" = "E6..Including.yourself..how.many.adults.and.children..live.in.your.household...The.individuals.who.have.been.residing.in.your.household.continuously.for.the.past.six.months"
    )

```


```{r}
# Set variable/column labels

   ssu_sece_clean <- ssu_sece_clean %>%
     set_variable_labels(.labels = variable_labels) # Requires package labelled

# ssu_sece_clean <- ssu_sece_clean %>%
#     set_variable_labels(value = variable_labels)


```


Using the function View() to open the new data in a spreadsheet reveals the column names and labels.

```{r}

View(ssu_sece_clean)

```



```{r}

# Drop un needed columns and save clean data

ssu_sec_e_arch <- ssu_sece_clean %>%
  select(c(1:2, 5:6, 9, 12:13))

# HINT: Passing View() to a pipeline produces NULL data structure.
# Avoid the function within a pipeline

```


## Generate a data dictionary

Generate a data dictionnary and search for variables with `look_for()`

Requires package labelled.



```{r}
# Data dictionary


ssu_B1sec_e_dict <- look_for(ssu_sec_e_arch, labels = TRUE, values = TRUE) 

# Fomart table

#flextable(data_dict)

#gt(data_dict)

kable(ssu_B1sec_e_dict)

```



## Export data to CSV


```{r}

#write.csv(ssu_B1sec_e_dict, "ssuB1_sec_e_arch.csv", row.names = FALSE)

```


